---
title: "ClimateVar_andFisheries_v1"
author: "Erica Ferrer"
date: "8/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)

# loading librarires
library(tidyverse)
library(lubridate)
library(ggthemes)
library(purrr)
library(patchwork)

# setting default plotting theme
theme_set(theme_tufte())

```


```{r}
# this code source a function to wrangle csv data into one dataset. 

source("01_data_wrangling.R")
```



## View/Explore the Environmental Data (supplied by Fabio)
```{r fig.cap='A) is the time series data of SST for the two regions with an hisorical average (black line) and a trend line (red curve); B) is the time series of SST anomalies calculated on the historical average for each region, vertical dashed lines correspond to El Nino events, the bold line is for the lastest El Nino event of 2014; C) is the time series data of Chl-a for the two regions with an hisorical average (black line) and a trend line (red curve); B) is the time series of Chl-a anomalies calculated on the historical average for each region, vertical dashed lines correspond to El Nino events, the bold line is for the lastest El Nino event of 2014.', fig.align='center', fig.height=8, fig.width=8}

# plotting SST ----------------------------------------------------------------

p1 <- env %>% 
        group_by(region) %>% 
        mutate(tot_mean= mean(sst, na.rm = T)) %>% 
        ggplot(aes(x=as.Date(date), y=sst, col=region))+
        geom_line(cex=.4)+
        labs(x="", y="SST °C", title = "A")+
        geom_hline(aes(yintercept=tot_mean), col = "black")+
        geom_smooth(col="red")+
        facet_grid(region~.,)+
        scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
        scale_color_brewer(palette = 2, aesthetics="black")+
        theme(legend.position = "", 
              panel.grid = element_blank(), 
              axis.text.x = element_text(angle=90))

## SST anomalies

p2 <- env %>% 
        group_by(region) %>% 
        mutate(month = format(date, "%m"), year = format(date, "%Y")) %>% 
        group_by(region, month) %>% 
        mutate(average = mean(sst, na.rm = T), anomaly = sst-average) %>% 
        ggplot(aes(x=as.Date(date), y=anomaly)) +
        geom_col(aes(fill= anomaly>0))+
        geom_hline(yintercept = 0)+
        geom_vline(xintercept=
                as.Date(c("2004-01-01 UTC", 
                          "2006-01-01 UTC", 
                          "2009-01-01 UTC", 
                          "2018-01-01 UTC")),
                   linetype=4, colour="black")+
        geom_vline(xintercept=as.Date("2014-01-01 UTC"),
                   linetype=1, colour="black")+
        scale_fill_manual(labels = c("FALSE" = "Lower than avg", 
                                     "TRUE"="Higher than avg"), 
                          values = c("blue", "red"))+
        labs(x ="", y = "SST anomaly", title = "B")+
        scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
        facet_grid(region~.,)+
        theme(legend.title = element_blank(), 
              panel.grid = element_blank(), 
              axis.text.x = element_text(angle=90))

# plotting Chl-a -----------------------------------------------------------------


p3 <- env %>% 
        group_by(region) %>% 
        mutate(tot_mean= mean(chla, na.rm = T)) %>% 
        ggplot(aes(x=as.Date(date), y=chla, col=region))+
        geom_line(cex=.4)+
        labs(x="", y="Chl-a concentration", title = "C")+
        geom_hline(aes(yintercept=tot_mean), col = "black")+
        geom_smooth(col="red")+
        facet_grid(region~.,)+
        ylim(0,10)+
        scale_color_brewer(palette = 2, aesthetics="black")+
        scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
        theme(legend.position = "", 
              panel.grid = element_blank(), 
              axis.text.x = element_text(angle=90))

# Chl-a Anomalies

p4 <- env %>% 
        group_by(region) %>% 
        mutate(month = format(date, "%m"), year = format(date, "%Y")) %>% 
        group_by(region, month) %>% 
        mutate(average = mean(chla, na.rm = T), anomaly = chla-average) %>% 
        ggplot(aes(x=as.Date(date), y=anomaly)) +
        geom_col(aes(fill= anomaly>0))+
        geom_hline(yintercept = 0)+
        geom_vline(xintercept=
                as.Date(c("2004-01-01 UTC", 
                          "2006-01-01 UTC", 
                          "2009-01-01 UTC", 
                          "2018-01-01 UTC")),
                   linetype=4, colour="black")+
        geom_vline(xintercept=as.Date("2014-01-01 UTC"),
                   linetype=1, colour="black")+
        scale_fill_manual(labels = c("FALSE" = "Lower than avg", 
                                     "TRUE"="Higher than avg"), 
                          values = c("black", "green"))+
        labs(x ="", y = "Chl-a anomaly", title = "D")+
        scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
        facet_grid(region~.,)+
        theme(legend.title = element_blank(), 
              panel.grid = element_blank(), 
              axis.text.x = element_text(angle=90))

# this just add plots together
p1+p2+p3+p4
```





## Interactive plots

**These are the same plots as above but are interactive so you can scroll and zoom.**

```{r PLOT: SST in the Upper Gulf & Baja Pacific}
plotly::ggplotly(p1)
```

```{r PLOT: SST Anomalies in the UG & BC-P}
plotly::ggplotly(p2)
```

```{r PLOT: Chl-a in the Upper Gulf & Baja Pacific}
plotly::ggplotly(p3)
```

```{r PLOT: Chl-a Anomalies in the UG & BC-P}
plotly::ggplotly(p4)
```



## Load in CONAPESCA Fisheries data
```{r Crate dataframe for Mexican pacific}
MexicoPacific_Data <- read.csv(file="data_pacifico_Ferrer_vCSV copy.csv", header=TRUE) #710882 obs.
#plyr::count(MexicoPacific_Data$State)

SSF_MexicoPacific_Data<- MexicoPacific_Data[(MexicoPacific_Data$Tipo.de.pesca=="ARTESANAL"),] #698604 obs.

names(SSF_MexicoPacific_Data) <- c("Year", "Month", "State", "Type", "LandingSite","GeneralName","ScientificName","TrophicLevel", "Catch_kg", "AvgPrice_MXN", "Total_MXN", "Total_USD")
```

```{r Create dataframe for Sonora}
# plyr::count(SSF_MexicoPacific_Data$State)
# BAJA CALIFORNIA= 60431; BAJA CALIFORNIA SUR= 139422; SINALOA= 98154; SONORA= 35902
# So, we would expect any subset the data from NW Mexico to incude 333909 records.

library(dplyr)
SSF_nw.Mexico<- SSF_MexicoPacific_Data %>%
        select(Year,Month, State, Type, LandingSite, GeneralName, ScientificName, TrophicLevel, Catch_kg, AvgPrice_MXN, Total_MXN, Total_USD) %>%
        filter(State=="BAJA CALIFORNIA" | State=="BAJA CALIFORNIA SUR" | State=="SINALOA" | State=="SONORA") # 333909 obs. Good!
#plyr::count(SSF_nw.Mexico$State)

# Fisheries dataframe for Sonora----------------------------------------------------------
SSF_Sonora<- filter(SSF_nw.Mexico, State=="SONORA") #35902 obs. Good!

# Environmental dataframe for Sonora------------------------------------------------------
## SST + Chl-a for the region approximately surrounding Sonora
MonthlyEnv<- env %>%
        filter(region=="UpperGulf")  # 204 obs. 4 variables
```

## FROM HERE ON OUT, ALL DATA REFERS to the SONORAN region (/upper GoC).
This analysis can be expanded and and made more precise + accurate later on.

```{r Abridge Sonoran dataframe}
SSF_Sonora<- SSF_Sonora %>%
        select(-c(Type, TrophicLevel, AvgPrice_MXN, Total_MXN, Total_USD))
```

```{r Remove functional groups that occur fewer than 30% of months- PART 1}
print(min(SSF_Sonora$Year))
print(max(SSF_Sonora$Year))
# 18 years total in time series. Remove species that do not occur at least 9 years in each state.

plyr::count(SSF_Sonora$ScientificName)

# Perhaps this should done at the species level rather the "common name" level?
# A question for another time. For now, proceed at the level of common names.

MonthlyCatches<- SSF_Sonora %>%
        group_by(Year, Month, State, GeneralName) %>%
        summarize(sum(Catch_kg))

# Quality Control Check: do these #s match?
print(sum(SSF_Sonora$Catch_kg))
print(sum(MonthlyCatches$`sum(Catch_kg)`))
# Yes!
```

```{r Remove functional groups that occur fewer than 30% of months- PART 2}
plyr::count(MonthlyCatches$GeneralName)

MonthlyCatches.f<- MonthlyCatches %>%
        group_by(GeneralName) %>%
        mutate(freq= n()) %>%
        ungroup %>%
        filter(freq>72) %>%     #why 72? There are 216 months total. 30% of 216 is 72.
        select(-freq)   # remove common name taxa that occur less than 30% of months

# NOTE ABOUT NAMING: ".f" is for "filtered" dataframe

# QC Check:
plyr::count(MonthlyCatches.f$GeneralName) #Good!
```

```{r Add month # to Sonoran monthly catches}
# Does anyone know an easier way to convert months listed in spanish?

MonthlyCatches.f$MonthNumber<- c(NA)

index= length(MonthlyCatches.f$MonthNumber)

for (i in 1:index){
        if (MonthlyCatches.f$Month[i]=="enero"){
                MonthlyCatches.f$MonthNumber[i]<- 1     
        } else if (MonthlyCatches.f$Month[i]=="febrero"){
                MonthlyCatches.f$MonthNumber[i]<- 2    
        } else if (MonthlyCatches.f$Month[i]=="marzo"){
                MonthlyCatches.f$MonthNumber[i]<- 3   
        } else if (MonthlyCatches.f$Month[i]=="abril"){
                MonthlyCatches.f$MonthNumber[i]<- 4
        } else if (MonthlyCatches.f$Month[i]=="mayo"){
                MonthlyCatches.f$MonthNumber[i]<- 5 
        } else if (MonthlyCatches.f$Month[i]=="junio"){
                MonthlyCatches.f$MonthNumber[i]<- 6
        } else if (MonthlyCatches.f$Month[i]=="julio"){
                MonthlyCatches.f$MonthNumber[i]<- 7
        } else if (MonthlyCatches.f$Month[i]=="agosto"){
                MonthlyCatches.f$MonthNumber[i]<- 8
        } else if (MonthlyCatches.f$Month[i]=="septiembre"){
                MonthlyCatches.f$MonthNumber[i]<- 9
        } else if (MonthlyCatches.f$Month[i]=="octubre"){
                MonthlyCatches.f$MonthNumber[i]<- 10
        } else if (MonthlyCatches.f$Month[i]=="noviembre"){
                MonthlyCatches.f$MonthNumber[i]<- 11
        } else if (MonthlyCatches.f$Month[i]=="diciembre"){
                MonthlyCatches.f$MonthNumber[i]<- 12  
        }
}#index

# Quality Control check: do these #s match?
#plyr::count(MonthlyCatches.f$Month)
#plyr::count(MonthlyCatches.f$MonthNumber)
# Yes they do!
```

## Principle Component Analysis (PC Analysis)
# Helpful guide to PCA: https://www.datacamp.com/community/tutorials/pca-analysis-r

```{r PC Analysis- PART 1: preparing the dataframes}
# Need to restructure data column such that month-year are the first column, and all subsequent columns are common names. Each cell will be the total catch for taxa X in any give year. But first, we'll need to restructure the dataframes a little bit...


# Fisheries MONTHLY dataframe (for PCA)---------------------------------

# 1st: discard columns that aren't necessary right now.
MonthlyCatches.f<- MonthlyCatches.f[,-c(2,3)]
# 2nd: rename columns
names(MonthlyCatches.f)<- c("Year", "GeneralName", "Catch_kg", "Month")
# 3rd: reorder columns
MonthlyCatches.f<- MonthlyCatches.f[c("Year","Month","GeneralName","Catch_kg")] # 5609 obs.

# 3rd: transpose data, spread "common names" out as columns (necessary to perform a PCA)
# Helpful guide to transposing: https://www.r-statistics.com/tag/transpose/
library(reshape)
monthly.PCA_df<- cast(MonthlyCatches.f, Year+Month~ GeneralName) #18 obs. 30 variables.

# Replace NAs with 1 (for now at least, because PCA doesn't do well with NAs or 0s)
monthly.PCA_df[is.na(monthly.PCA_df)] <- 1


# Fisheries ANNUAL dataframe (for PCA)---------------------------------

#Summarize data by year to run ANNUAL PCAs
AnnualCatches.f<- MonthlyCatches.f %>%
        group_by(Year,GeneralName) %>%
        summarize(sum(Catch_kg)) #521 obs.

names(AnnualCatches.f) <- c("Year", "GeneralName","AnnualCatch_kg") # 521 obs.

# Quality Control Check: do these #s match?
#print(sum(MonthlyCatches.f$`sum(Catch_kg)`))
#print(sum(AnnualCatches.f$AnnualCatch_kg))
# Yes!
```

```{r Now prepare the SST and Chl-a data}
# BREAK UP THE DATE COLUMN using lubridate
MonthlyEnv$date <- as.Date(MonthlyEnv$date, format="%m-%d-%y")
library(lubridate)
MonthlyEnv$year <- year(ymd(MonthlyEnv$date))
MonthlyEnv$month <- month(ymd(MonthlyEnv$date)) 

# Re-structure MonthlyEnv for ease of use:
# 1st: discard columns
MonthlyEnv<- MonthlyEnv[,-c(1,2)]
# 2nd: rename columns
names(MonthlyEnv)<- c("mean_Chla", "mean_SST", "Year", "Month")
# 3rd: reorder columns
MonthlyEnv<- MonthlyEnv[c("Year", "Month", "mean_SST", "mean_Chla")]

# Enviro MONTHLY dataframe (for PCA)---------------------------------
#MonthlyEnv

# Enviro ANNUAL dataframe (for PCA)---------------------------------
AnnualEnv<- MonthlyEnv %>%
        group_by(Year) %>%
        summarise(mean(mean_SST), mean(mean_Chla), sd(mean_SST), sd(mean_Chla))

names(AnnualEnv)<- c("Year", "mean_SST", "mean_Chla", "sd_SST", "sd_Chla")
        
plot(AnnualEnv$Year, AnnualEnv$mean_SST, type="l")
plot(AnnualEnv$Year, AnnualEnv$sd_SST, type="l")
```

```{r PC Analysis- PART 2: Enviro PCAs}
# PCA for MONTHLY change in enviro parameters off Sonora----------------------------------

## Prepare Enviro dataset:
MonthlyEnv<- MonthlyEnv[-c(204),] #remove row 204 due to NA

## RUN PCA
enviroPCA_monthly <- prcomp(MonthlyEnv[,c(3,4)], center = TRUE,scale. = TRUE)
summary(enviroPCA_monthly)
# 82.7% of variance explained by first PC

# Extract enviro PC 1 from this analysis:
enviroPC1_monthly<- enviroPCA_monthly$x[,'PC1']

# Store PC 1 as a dataframe for ease of reference:
enviroPC1_monthly<- data.frame(enviroPC1_monthly)
enviroPC1_monthly<- cbind(enviroPC1_monthly, MonthlyEnv$Year, MonthlyEnv$Month)
names(enviroPC1_monthly) <- c("enviroPC1_monthly", "Year", "Month")


# PCA for ANNUAL change in enviro parameters off Sonora----------------------------------

## Prepare Enviro dataset:
AnnualEnv<- AnnualEnv[-c(17),] #remove row 17, year 2019, due to NAs.
names(AnnualEnv) <- c("Year", "mean_SST", "mean_Chla", "SST_sd", "Chla_sd")

## RUN PCA
enviroPCA_annual <- prcomp(AnnualEnv[,c(3,4)], center = TRUE,scale. = TRUE)
summary(enviroPCA_annual)
# 73.0% of variance explained by first PC

# Extract enviro PC 1 from this analysis:
enviroPC1_annual<- enviroPCA_annual$x[,'PC1']

# Give enviro PC 1 a dataframe for reference:
enviroPC1_annual<- data.frame(enviroPC1_annual)
enviroPC1_annual<- cbind(enviroPC1_annual, AnnualEnv$Year)
names(enviroPC1_annual) <- c("enviroPC1_annual", "Year")
```



```{r PC Analysis- PART 3: Fishy PCAs}
# PCA for ANNUAL change in fish communities off Sonora----------------------------------

# Prepare dataset:
# Replace NAs with 1 (for now at least, because PCA doesn't do well with NAs or 0s)
#monthly.PCA_df[is.na(monthly.PCA_df)] <- 1

## RUN PCA:
fishyPCA_monthly <- prcomp(monthly.PCA_df[,c(3:31)], center = TRUE,scale. = TRUE)
summary(fishyPCA_monthly)
# 15.6% of variance explained by PC1
# 24.6% of var. explained by fist two PCs

# Extract fishy PC1 and PC2 from this analysis:
fishyPC1_monthly<- fishyPCA_monthly$x[,'PC1']

# Give fishy PC 1 (monthly) a dataframe for reference:
fishyPC1_monthly<- data.frame(fishyPC1_monthly)
fishyPC1_monthly<- cbind(fishyPC1_monthly, monthly.PCA_df$Year, monthly.PCA_df$Month) #216 obs. Years 2001-2018
names(fishyPC1_monthly) <- c("fishyPC1_monthly", "Year", "Month")


# PCA for ANNUAL change in fish communities off Sonora----------------------------------

# Like before, we must transpose this dataframe:
annual.PCA_df<- cast(AnnualCatches.f, Year~ GeneralName) #18 obs. 30 variables.

# One NA present. Rubia y Villajaiba in year 2005... ??? Exclude for now. 
annual.PCA_df<- annual.PCA_df[-c(28)]
# COME BACK TO THIS LATER. In the future, we could fill this NA with a "1" to run the PCA.

# RUN PCA:
fishyPCA_annual <- prcomp(annual.PCA_df[,c(2:29)], center = TRUE,scale. = TRUE)

summary(fishyPCA_annual)
# 32.5% of variance explained by first PC
# 47.2% of variance explained by first two PCs
# 57.5% of variance explained by first three PCs

str(fishyPCA_annual)

# Extract fishy PC 1 from this analysis:
fishyPC1_annual<- fishyPCA_annual$x[,'PC1']

# Give fishy PC 1 (annual) a dataframe for reference:
fishyPC1_annual<- data.frame(fishyPC1_annual)
fishyPC1_annual<- cbind(fishyPC1_annual, annual.PCA_df$Year) #18 obs. Years 2001-2018
names(fishyPC1_annual) <- c("fishyPC1_annual", "Year")


## PCA plots

# Eigenvalues for the PCA
library("factoextra")
eig.val <- fishyPCA_annual$sdev^2
eig.val

# Scree plot for PCA
fviz_eig(fishyPCA_annual, addlabels = TRUE, ylim = c(0, 50))

#library(ggbiplot)
#ggbiplot(pcobj= Sonora.PCA, obs.scale = 1, var.scale = 1)

library(ggfortify)
autoplot(fishyPCA_annual, data = annual.PCA_df, colour = 'Year')
```

```{r Merge dataframes together as appropriate}
## Prepare dataframes 
#Check that lengths of all dataframes are equal
length(monthly.PCA_df$Year) #216. Years 2001-2018
length(fishyPC1_monthly$Year) #216. Years 2001-2018
length(enviroPC1_monthly$Year) #203. Years 2003-(november) 2019
length(MonthlyEnv$Year) # 203 obs. Years 2003-(november) 2019

# Need to trim these dataframes to fit eachother
monthly.PCA_df<-monthly.PCA_df[-c(1:24),] # 192 obs.
fishyPC1_monthly<-fishyPC1_monthly[-c(1:24),] # 192 obs.
enviroPC1_monthly<-enviroPC1_monthly[-c(193:203),] # 192 obs.
MonthlyEnv<-MonthlyEnv[-c(193:203),] # 192 obs.

# The same is also true for the ANNUAL dataframes
annual.PCA_df<- annual.PCA_df[-c(1,2),] # 16 obs. 29 variables. Removes years 2001 & 2002.
fishyPC1_annual<- fishyPC1_annual[-c(1,2),] # 16 obs. Removes years 2001 & 2002.
length(enviroPC1_annual$enviroPC1_annual) # 16 obs.

# Final dataframe for MONTHLY data off Sonora----------------------------------
## BEGIN MERGE of MONTHLY dataframes
monthlyFish.Env_df<- merge(monthly.PCA_df, MonthlyEnv, by=c("Year","Month"), sort = TRUE) #192 obs of 33 variables. Years 2003-2018
# ADD fishy PC 1
monthlyFish.Env_df<- merge(monthlyFish.Env_df, fishyPC1_monthly, by=c("Year","Month"), sort = TRUE)
# ADD enviro PC 1
monthlyFish.Env_df<- merge(monthlyFish.Env_df, enviroPC1_monthly, by=c("Year","Month"), sort = TRUE)

# Final dataframe for ANNUAL data off Sonora----------------------------------
## BEGIN MERGE of ANNUAL dataframes
annualFish.Env_df<- merge(annual.PCA_df, AnnualEnv, by=c("Year"), sort = TRUE) #16 obs of 33 variables. Years 2003-2018
# ADD fishy PC 1
annualFish.Env_df<- merge(annualFish.Env_df, fishyPC1_annual, by=c("Year"), sort = TRUE)
# ADD enviro PC 1
annualFish.Env_df<- merge(annualFish.Env_df, enviroPC1_annual, by=c("Year"), sort = TRUE)
```



## Create a CORRELATION MATRIX for fisheries/enviro parameters with significance levels

```{r Correlation matrix of ANNUAL fisheries-enviro with significance levels}
# Correlation matrix with significance levels
res.annual <- cor(annualFish.Env_df)

library("Hmisc")
res2.annual <- rcorr(as.matrix(annualFish.Env_df))
#res2$r
#res2$n
#res2$P
#res2
```

```{r Simplify the correlation matrix: visually exclude some fisheries}
# Think about which taxa we are most interested in at the moment: shrimp, crab, lobsters, clams, snails, sharks, rays, and a sampling of fin fish species. 
# For now, I just want to narrow down the picture a bit to help tell a story.

# Exclude these taxa:
annualFish.Env_df.f<- annualFish.Env_df[,-which(names(annualFish.Env_df) %in% c("BAGRE", "BANDERA","CARPA", "CAZON", "LANGOSTA", "MERO","PAMPANO", "PIERNA", "RONCO", "OTRAS"))]

# Reorder columns for better presentation:
order.vector <- c("Year", "enviroPC1_annual", "mean_SST", "SST_sd", "mean_Chla", "Chla_sd", "fishyPC1_annual", "ALMEJA", "CARACOL", "CAMARON", "JAIBA", "CALAMAR", "PULPO", "BERRUGATA", "CABRILLA", "CORVINA", "GUACHINANGO", "PARGO", "LENGUADO", "MOJARRA", "LISA", "JUREL", "RAYA Y SIMILARES", "SIERRA", "TIBURON")
cnames <- order.vector[order.vector %in% names(annualFish.Env_df.f)]

annualFish.Env_df.f<- annualFish.Env_df.f[cnames]

# Again, correlation matrix:
res.annual <- cor(annualFish.Env_df.f)

library("Hmisc")
res2.annual <- rcorr(as.matrix(annualFish.Env_df.f))
res2.annual
```


```{r CORRELOGRAM: for a simplified correlation matrix}
# Helpful info on how to make a heatmap: http://www.sthda.com/english/wiki/correlation-matrix-how-to-make-a-heatmap
# Helpful on how to make a correlogram: http://www.sthda.com/english/wiki/visualize-correlation-matrix-using-correlogram#visualization-methods

cor.mtest <- function(mat, ...) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(mat[, i], mat[, j], ...)
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}

# matrix of the p-value of the correlation
p.mat <- cor.mtest(annualFish.Env_df.f)
head(p.mat)

library(corrplot)
library(RColorBrewer)
corrplot(res.annual, type = "upper", order = "original", 
         tl.col = "black", tl.srt = 30, tl.cex=.5, diag= FALSE, col= brewer.pal(n=8, name="RdYlBu"), p.mat = p.mat, sig.level = 0.15, insig = "blank" , title= "Correlogram for the Fisheries of Sonora, p< 0.15", mar=c(0,0,1,0))
```

```{r CORRELOGRAM: Invertebrate Fisheries with significance level}
Inverts.Env_df<- annualFish.Env_df.f[,-c(14:25)]

res.Inverts <- cor(Inverts.Env_df)

# matrix of the p-value of the correlation
p.mat.Inverts <- cor.mtest(Inverts.Env_df)
#head(p.mat.Inverts)

library(corrplot)
library(RColorBrewer)
corrplot(res.Inverts, type = "upper", order = "original", 
         tl.col = "black", tl.srt = 30, tl.cex=.5, diag= FALSE, col= brewer.pal(n=8, name="RdYlBu"), p.mat = p.mat.Inverts, insig = "blank", title= "Invertebrate Correlogram, p< 0.15", mar=c(0,0,1,0))
```

```{r Who are the "Caracols" of Sonora?}
Caracols<- SSF_Sonora[SSF_Sonora$GeneralName=="CARACOL",]
plyr::count(Caracols$ScientificName)
```

```{r Who are the "Carpas" & "Lisas" of Sonora?}
Carpa<- SSF_Sonora[SSF_Sonora$GeneralName=="CARPA",]
plyr::count(Carpa$ScientificName)
# The CARPAS have been excluded from the fisheries data since they are freshwater fish

Lisa<- SSF_Sonora[SSF_Sonora$GeneralName=="LISA",]
plyr::count(Lisa$ScientificName)
```



```{r PLOT: fisheries landings over time}
library(ggplot2)

plot(Inverts.Env_df$Year, Inverts.Env_df$CARACOL)

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

p5 <- Inverts.Env_df %>% 
        select(CARACOL, JAIBA, CAMARON, CALAMAR, ALMEJA, Year) %>%
        ggplot()+
        geom_line(aes(x=Year, y=c(CARACOL), col="CARACOL")) +
        geom_line(aes(x=Year, y=c(JAIBA), col="JAIBA")) +
        geom_line(aes(x=Year, y=c(CAMARON), col="CAMARON")) +
        geom_line(aes(x=Year, y=c(CALAMAR), col="CALAMAR")) +
        geom_line(aes(x=Year, y=c(ALMEJA), col="ALMEJA")) +
        ggtitle("Fisheries Landings from 2003-2018") +
        xlab("Year") +
        scale_y_continuous(name = "Landings (in kg)") +
        theme(text=element_text(family="Helvetica"), plot.title = element_text(size=10), legend.position= "none")+
        theme(axis.line.x = element_line(color="black", size = .3),
              axis.line.y = element_line(color="black", size = .3)) +
        geom_vline(aes(xintercept=2009), col = "grey", linetype="dotted")

p5

# Add a second axis and specify its features
#sec.axis = sec_axis( trans=~.*10, name="Second Axis")


p6 <- Inverts.Env_df %>% 
        select(CARACOL, JAIBA, CAMARON, CALAMAR, ALMEJA, Year) %>%
        filter(Year>2008) %>%
        ggplot()+
        geom_line(aes(x=Year, y=c(CARACOL), col="GASTROPOD")) +
        geom_line(aes(x=Year, y=c(JAIBA), col="CRAB")) +
        geom_line(aes(x=Year, y=c(CAMARON), col="SHRIMP")) +
        geom_line(aes(x=Year, y=c(CALAMAR), col="SQUID")) +
        geom_line(aes(x=Year, y=c(ALMEJA), col="CLAM")) +
        ggtitle("Fisheries Landings since 2009") +
        xlab("Year") +
        ylab("Landings (in kg)") +
        theme(text=element_text(family="Helvetica"), plot.title = element_text(size=10))+
        theme(axis.line.x = element_line(color="black", size = .3),
              axis.line.y = element_line(color="black", size = .3))

p5+p6


# other default fonts include: sans, serif, Courier, Helvetica, Times, AvantGarde, Bookman, Helvetica-Narrow, NewCenturySchoolbook, Palatino, URWGothic, URWBookman, NimbusMon, URWHelvetica, NimbusSan, NimbusSanCond, CenturySch, URWPalladio, URWTimes, NimbusRom
# See here for more about fonts: https://stackoverflow.com/questions/34522732/changing-fonts-in-ggplot2
```

```{r PLOT: Fishy PC versus mean SST and mean Chl-a}
#Helpful guide on applying a second axis in ggplot: https://www.r-graph-gallery.com/line-chart-dual-Y-axis-ggplot2.html
# Although I won't use this just yet.

p7 <- annualFish.Env_df.f %>% 
        select(fishyPC1_annual, Year) %>%
        #filter(Year>2008) %>%
        ggplot()+
        geom_line(aes(x=Year, y=c(fishyPC1_annual)), col="red") +
        ggtitle("Fishy PC 1 over time") +
        xlab("Year") +
        ylab("PC: trend in make-up fisheries landings") +
        theme(text=element_text(family="Helvetica"))+
        theme(axis.line.x = element_line(color="black", size = .3),
              axis.line.y = element_line(color="black", size = .3))

p7


p8 <- annualFish.Env_df.f %>% 
        select(mean_SST, Year) %>%
        #filter(Year>2008) %>%
        ggplot()+
        geom_line(aes(x=Year, y=c(mean_SST)), color="blue") +
        ggtitle("SST over time") +
        xlab("Year") +
        ylab("Temperature (C)") +
        theme(text=element_text(family="Helvetica"))+
        theme(axis.line.x = element_line(color="black", size = .3),
              axis.line.y = element_line(color="black", size = .3))

p8

p7+p8
```

```{r NW Mexico fisheries landings and revenue by class}
plyr::count(SSF_MexicoPacific_Data$GeneralName)

PacificSpeciesData<- read.csv(file="MexPacificSpecies_CommonNames.csv", header=TRUE)
plyr::count(PacificSpeciesData$GeneralName)

Test<- merge(PacificSpeciesData, SSF_nw.Mexico, by=c("GeneralName"), sort = TRUE)

Test2<- Test[,c(1,2,3,4,5,10,13)]

Test2<- Test %>%
        group_by(State, Year, Class) %>%
        summarise(sum(Catch_kg), sum(Total_USD))

# Landings + value SINCE 2014
Test2<- Test2 %>% 
        filter(Year>2013)

Test2<- Test2[!(Test2$Class=="Exclude"),]

Test3<- Test2 %>%
        group_by(State, Class) %>%
        summarise(sum(`sum(Catch_kg)`), sum(`sum(Total_USD)`))

#plyr::count(Test3$Class)

names(Test3)<- c("State", "Class", "TotalCatch_kg", "TotalValue_USD")

## PIE CHARTS (landings + value since 2014)
slices <- Test3[(1:10),]$TotalCatch_kg
lbls <- c("Algae", "Cephalopods", "Crustaceans", "DemersalMolluscs", "Echinoderm", "LargePelagics", "NonResDemFish", "ResidentDemFish", "Shrimp", "SmallPelagics")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
BajaCatch_pie<- pie(slices,labels = lbls, col=rainbow(length(lbls)),
   main="SSF Catch Composition: Baja California")

slices <- Test3[(11:20),]$TotalCatch_kg
lbls <- c("Algae", "Cephalopods", "Crustaceans", "DemersalMolluscs", "Echinoderm", "LargePelagics", "NonResDemFish", "ResidentDemFish", "Shrimp", "SmallPelagics")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
BCSCatch_pie<- pie(slices, labels = lbls, col=rainbow(length(lbls)),
   main="SSF Catch Composition: Baja California Sur")

slices <- Test3[(21:30),]$TotalCatch_kg
lbls <- c("Algae", "Cephalopods", "Crustaceans", "DemersalMolluscs", "Echinoderm", "LargePelagics", "NonResDemFish", "ResidentDemFish", "Shrimp", "SmallPelagics")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
SinaloaCatch_pie<- pie(slices, labels = lbls, col=rainbow(length(lbls)),
   main="SSF Catch Composition: Sinaloa")


New <- data.frame("SONORA", "Algae", 0, 0)
Test3[nrow(Test3)+1,] <- NA
Test3[40,]<- New
Test3 <- Test3[c(1:30,40,31:39),]

slices <- Test3[(31:40),]$TotalCatch_kg
lbls <- c("Algae", "Cephalopods", "Crustaceans", "DemersalMolluscs", "Echinoderm", "LargePelagics", "NonResDemFish", "ResidentDemFish", "Shrimp", "SmallPelagics")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
SonoraCatch_pie<- pie(slices, labels = lbls, col=rainbow(length(lbls)),
   main="SSF Catch Composition: Sonora")

Test4<- Test3 %>%
        group_by(Class) %>%
        summarise(sum(TotalCatch_kg), sum(TotalValue_USD))
names(Test4)<- c("Class", "TotalCatch_kg", "TotalValue_USD")

slices <- Test4$TotalCatch_kg
lbls <- c("Algae", "Cephalopods", "Crustaceans", "DemersalMolluscs", "Echinoderm", "LargePelagics", "NonResDemFish", "ResidentDemFish", "Shrimp", "SmallPelagics")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
NWmexicoCatch_pie<- pie(slices, labels = lbls, col=rainbow(length(lbls)),
   main="SSF Catch Composition: NW Mexico")

slices <- Test4$TotalValue_USD
lbls <- c("Algae", "Cephalopods", "Crustaceans", "DemersalMolluscs", "Echinoderm", "LargePelagics", "NonResDemFish", "ResidentDemFish", "Shrimp", "SmallPelagics")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
NWmexicoRevenue_pie<- pie(slices, labels = lbls, col=rainbow(length(lbls)),
   main="SSF Revenue Composition: NW Mexico")
```

```{r}
slices <- Test3[(1:10),]$TotalValue_USD
lbls <- c("Algae", "Cephalopods", "Crustaceans", "DemersalMolluscs", "Echinoderm", "LargePelagics", "NonResDemFish", "ResidentDemFish", "Shrimp", "SmallPelagics")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
SonoraCatch_pie<- pie(slices, labels = lbls, col=rainbow(length(lbls)),
   main="SSF Revenue Composition: Baja California")

slices <- Test3[(11:20),]$TotalValue_USD
lbls <- c("Algae", "Cephalopods", "Crustaceans", "DemersalMolluscs", "Echinoderm", "LargePelagics", "NonResDemFish", "ResidentDemFish", "Shrimp", "SmallPelagics")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
SonoraCatch_pie<- pie(slices, labels = lbls, col=rainbow(length(lbls)),
   main="SSF Revenue Composition: Baja California Sur")

slices <- Test3[(21:30),]$TotalValue_USD
lbls <- c("Algae", "Cephalopods", "Crustaceans", "DemersalMolluscs", "Echinoderm", "LargePelagics", "NonResDemFish", "ResidentDemFish", "Shrimp", "SmallPelagics")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
SonoraCatch_pie<- pie(slices, labels = lbls, col=rainbow(length(lbls)),
   main="SSF Revenue Composition: Sinaloa")

slices <- Test3[(31:40),]$TotalValue_USD
lbls <- c("Algae", "Cephalopods", "Crustaceans", "DemersalMolluscs", "Echinoderm", "LargePelagics", "NonResDemFish", "ResidentDemFish", "Shrimp", "SmallPelagics")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
SonoraCatch_pie<- pie(slices, labels = lbls, col=rainbow(length(lbls)),
   main="SSF Revenue Composition: Sonora")
```

```{r TESTING: breakdown of fisheries catch by specific species}
plyr::count(SSF_nw.Mexico$ScientificName[which(SSF_nw.Mexico$GeneralName=="JAIBA" & SSF_nw.Mexico$State=="SONORA")])

Test5<- SSF_nw.Mexico[which((SSF_nw.Mexico$ScientificName=="Farfantepenaeus californiensis" | SSF_nw.Mexico$ScientificName=="Litopenaeus stylirostris") & (SSF_nw.Mexico$State=="BAJA CALIFORNIA SUR")),]

Test5.2<- SSF_nw.Mexico[which(SSF_nw.Mexico$ScientificName=="Callinectes bellicosus" | SSF_nw.Mexico$ScientificName=="Callinectes sp" | SSF_nw.Mexico$ScientificName=="Brachyura"),]

Test5.3<- SSF_nw.Mexico[which(SSF_nw.Mexico$ScientificName=="Panopea generosa"),]

Test6<- Test5 %>%
        group_by(Year, ScientificName) %>%
        summarise(sum(Catch_kg))

Test6.2<- Test5.2 %>%
        group_by(Year, State) %>%
        summarise(sum(Catch_kg))

Test6.3<- Test5.3 %>%
        group_by(Year, State) %>%
        summarise(sum(Catch_kg))

plot(Test6$Year[which(Test6$ScientificName=="Farfantepenaeus californiensis")], Test6$`sum(Catch_kg)`[which(Test6$ScientificName=="Farfantepenaeus californiensis")], type="l")
abline(v = 2006, col = 'coral2', lwd = 2)
abline(v = 2008, col = 'blue', lwd = 2)
abline(v = 2010, col = 'coral2', lwd = 2)
# Shrimp up in Baja (starts in ~2006), Even in BCS (starts in ~2003), Even in Sonora 
# Large range
# Higher in the foodweb

plot(Test6$Year[which(Test6$ScientificName=="Litopenaeus stylirostris")], Test6$`sum(Catch_kg)`[which(Test6$ScientificName=="Litopenaeus stylirostris")], type="l")
abline(v = 2006, col = 'coral2', lwd = 2)
abline(v = 2008, col = 'blue', lwd = 2)
abline(v = 2010, col = 'coral2', lwd = 2)
# Noticeably down in Baja (beginning ~2008), Even in BCS ?, Even/Up in Sonora
# Smaller range
# Lower in the foodweb

plot(Test6.2$Year[which(Test6.2$State=="SONORA")], Test6.2$`sum(Catch_kg)`[which(Test6.2$State=="SONORA")], type="l")
# Up

plot(Test6.2$Year[which(Test6.2$State=="BAJA CALIFORNIA")], Test6.2$`sum(Catch_kg)`[which(Test6.2$State=="BAJA CALIFORNIA")], type="l")
# Down

plot(Test6.2$Year[which(Test6.2$State=="BAJA CALIFORNIA SUR")], Test6.2$`sum(Catch_kg)`[which(Test6.2$State=="BAJA CALIFORNIA SUR")], type="l", main ="Crabs in the GoC")
abline(v = 2003, col = 'coral2', lwd = 2)
abline(v = 2007, col = 'coral2', lwd = 2)
abline(v = 2010, col = 'coral2', lwd = 2)
abline(v = 2000, col = 'blue', lwd = 2)
abline(v = 2008, col = 'blue', lwd = 2)
abline(v = 2011, col = 'blue', lwd = 2)
abline(v = 2012, col = 'blue', lwd = 2)
# Crabs going up (kind of)

plot(Test6.3$Year[which(Test6.3$State=="BAJA CALIFORNIA SUR")], Test6.3$`sum(Catch_kg)`[which(Test6.3$State=="BAJA CALIFORNIA SUR")], type="l", main ="Clams in the GoC")
# Clams going up (kind of)




SSF_nw.Mexico.f<- SSF_nw.Mexico[which(SSF_nw.Mexico$Year==2014),]
#SSF_nw.Mexico.f<- SSF_nw.Mexico.f %>%
        #filter(State== "BAJA CALIFORNIA" | State=="BAJA CALIFORNIA SUR" | State=="SONORA")
# Exclude Sinaloa or not?

Test7<- sum(SSF_nw.Mexico.f$Catch_kg[which(SSF_nw.Mexico.f$ScientificName=="Litopenaeus stylirostris" | SSF_nw.Mexico.f$ScientificName=="Farfantepenaeus californiensis" | SSF_nw.Mexico.f$ScientificName=="Callinectes sp" | SSF_nw.Mexico.f$ScientificName=="Callinectes bellicosus" | SSF_nw.Mexico.f$ScientificName=="Panopea generosa" | SSF_nw.Mexico.f$ScientificName=="Megapitaria squalida")])/sum(SSF_nw.Mexico.f$Catch_kg)
# Years 2014-2018: 21.7% catch w Sinaloa, 13.9% catch w/o Sinaloa
# In 2014: 20.0% of catch (incl. Sinaloa), 12.9% of catch w/o Sinaloa

Test7.2<- sum(SSF_nw.Mexico.f$Total_USD[which(SSF_nw.Mexico.f$ScientificName=="Litopenaeus stylirostris" | SSF_nw.Mexico.f$ScientificName=="Farfantepenaeus californiensis" | SSF_nw.Mexico.f$ScientificName=="Callinectes sp" | SSF_nw.Mexico.f$ScientificName=="Callinectes bellicosus" | SSF_nw.Mexico.f$ScientificName=="Panopea generosa" | SSF_nw.Mexico.f$ScientificName=="Megapitaria squalida")])/sum(SSF_nw.Mexico.f$Total_USD)
# Years 2014-2018: 31.1% revenue, 43.8% revenue w Sinaloa (years 2014-2018)
# In 2014: 52% of revenue (incl. Sonora), 40% of revenue w/o Sinaloa


Mexico_Data <- read.csv(file="AllMexico_CONAPESCAdata_vCSV copy.csv", header=TRUE)
Mexico_Data<- Mexico_Data[(Mexico_Data$TipoPesca=="ARTESANAL"),]

NW.Mexico_Data <- Mexico_Data %>%
        filter(State=="BAJA CALIFORNIA" | State=="BAJA CALIFORNIA SUR" | State=="SONORA" | State=="SINALOA")

curious<- sum(NW.Mexico_Data$Catch_kg[which(NW.Mexico_Data$Year==2014)])/sum(Mexico_Data$Catch_kg[which(Mexico_Data$Year==2014)])
# 77% of ALL catch landed in NW Mexico in 2014 (incl. industrial); 56% w/o Sinaloa
# 54% of all SSF catch in 2014; 43% w/o Sinoloa
curious2<- sum(NW.Mexico_Data$Value_MXN[which(NW.Mexico_Data$Year==2014)])/sum(Mexico_Data$Value_MXN[which(Mexico_Data$Year==2014)])
# 45% total SSF revenue
```